---
title: "pml project"
author: "ju-ok"
date: "11/12/2020"
output: 
    html_document:
        keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
Details:  http://groupware.les.inf.puc-rio.br/har

## Data
The training data for this project can be accessed from:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv \
The test data are also available at:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

# Exploratory Data analysis
```{r eda}
# Load packages 
library(caret)

# Load the data sets 
training <- read.csv("./pml-training.csv")
testing <- read.csv("./pml-testing.csv")

# Explore data sets
# structure
str(training)

# dimensions
dim(training)
```

# Clean the data
```{r clean}
# Clean the data sets 

# remove columns with undesired info
training <- training[, -c(1:7)]
testing <- testing[, -c(1:7)]

# remove columns with NA        
training <- training[, colSums(is.na(training)) == 0]
testing <- testing[, colSums(is.na(testing)) == 0]

# remove all character columns except classe
training <- training[, sapply(training[ , -86], is.numeric)]
testing <- testing[, sapply(testing[ , -86], is.numeric)]

# dimensions of the training and test data
dim(training); dim(testing)
```

# Split the training data
The training data is split into 70% training set and 30% testing set.
```{r split}
# Split training data into trainingset and testingset
set.seed(1222)

inTrain <- createDataPartition(y = training$classe, 
                               p = 0.7,
                               list = FALSE)

trainingset <- training[inTrain, ]
testingset <- training[-inTrain, ]
```

# Fit models
## classification tree
```{r tree1}
# classification tree
modrpart <- train(classe ~., method = "rpart",  data = trainingset)

# plot the final model
plot(modrpart$finalModel, main = "Classification tree")
text(modrpart$finalModel, use.n = TRUE, all = TRUE, cex = 0.8)
```

```{r tree2}
# predict using testingset
predrpart <- predict(modrpart, testingset)

# evaluate the model
confusionMatrix(predrpart, as.factor(testingset$classe))
```

## Generalized boosted model
```{r gbm}
# gbm
control <- trainControl(method = "repeatedcv", number = 3, repeats = 1)
modgbm <- train(classe ~., method = "gbm",  data = trainingset, 
                trControl = control, verbose = FALSE)
```

```{r gbm2}
# predict using testingset
predgbm <- predict(modgbm, testingset)

# evaluate
confusionMatrix(predgbm, as.factor(testingset$classe))
```

## random forest
```{r rf}
# rf
modrf <- train(classe ~., method = "rf", data = trainingset,
               trControl = trainControl(method = "cv"), number = 3)
```

```{r rf2}
# predict using testingset
predrf <- predict(modrf, testingset)

# evaluate
confusionMatrix(predrf, as.factor(testingset$classe))
```

# Select and use best model
Based on the accuracy scores from the confusionMatrix, the random forest model has the highest accuracy followed by the generalized boosted model and classification tree respectively. The random forest model will be used on the `testing` data.
```{r final}
predfinal <- predict(modrf, testing)
predfinal
```
